\fontfamily{phv}\selectfont{Causal Discovery methods for Effective Connectivity}

### 2013-12-10 docs/paper/paper.tex
small comment addition in discussion.
heel wat plaatjes en toegevoegd, met wat tekst er omheen. Plaatsing van de plaatjes door latex is nog lelijk.

\section{Results}
At first, we have applied modified PC on the strucural connectivity data of six subjects.
To achieve stability, we have run the algorithm multiple times, in which the data was permuted randomly in each iteration.
The avarage of all iterations was used as a measure of certainty, i.e. how consistent the algorithm is for each point of data.

The results from the altorithms are presented in the form of matrices.
A connection from region $n$ to region $m$ is represented by point $(n,m)$ or $n,m$ in the matrix.
For strucural data, this matrix is therefore symmetrical.
In the directional data, point $(n,m)$ denotes a directed edge from $n$ to $m$.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum. %TODO left or right hemisphere?

Modified PC and EMS-PC do not differ in finding the structure.
The results are presented in figure ~\ref{fig:struct_full_mean} as the avarage over all six subjects, with 300 iterations per subject.
\begin{figure}[h!]
  \centering
  \includegraphics{images/struct_full_mean_gray}
  \caption{The avarage structure of all subjects}
  \label{fig:struct_full_mean}
\end{figure}
In this matrix, higher values represent strong certainties of the presence of a connection.

\begin{figure}[h!]
  \centering
  \includegraphics{images/structure_max}
  \caption{The avarage structure of all subjects, according to the DWI-measurements of Hinne et al.}
  \label{fig:struct_mean_max}
\end{figure}

One notable difference between the results found in Hinne et al. are that connections between the the two hemispheres are clearly visible as two diagonal lines. Connections within the hemispheres are less clearly visible, as well as connections within the cerebellum.

% Mod-PC

%

%

%

%

>>>>

\section{Results}
At first, we have applied modified PC on the structural connectivity data of six subjects.
To achieve stability, we have run the algorithm multiple times, in which the data was permuted randomly in each iteration.
The avarage of all iterations was used as a measure of certainty, i.e. how consistent the algorithm is for each point of data.

The results from the altorithms are presented in the form of matrices.
A connection from region $n$ to region $m$ is represented by point $(n,m)$ or $n,m$ in the matrix.
For strucural data, this matrix is therefore symmetrical.
In the directional data, point $(n,m)$ denotes a directed edge from $n$ to $m$.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum. %TODO left or right hemisphere?
Modified PC and EMS-PC do not differ in finding the structure.
The results are presented in figure ~\ref{fig:struct_avg} as the avarage over all six subjects, with 300 iterations per subject.
The results of Hinne et al. are shown in ~\ref{fig:struct_max} for comparison.

\begin{figure}[h!]
  \centering
  \includegraphics{images/struct_full_mean_gray}
  \caption{The avarage structure of all subjects, found with the PC-algorithm}
  \label{fig:struct_avg}
\end{figure}
In this matrix, higher values represent strong certainties of the presence of a connection.

\begin{figure}[h!]
  \centering
  \includegraphics{images/structure_max}
  \caption{The avarage structure of all subjects, according to the DWI-measurements of Hinne et al.}
  \label{fig:struct_max}
\end{figure}

One notable difference between the results found in Hinne et al. is the strong connections between homological areas, visible as two diagonal lines.
This is a known shortcoming of diffusion weighted imaging, as these connections are typically too long to measure correctly.
These connections most likely correspond to the corpus callosum, as each region is connected with its counterpart in the other hemisphere.
Connections within the hemispheres are less clearly visible, as well as connections within the cerebellum.

The second step of the PC-algorithm finds directions based on this structure.
In Figures ~\ref{fig:pdag_avg_mod} and ~\ref{fig:pdag_avg_ems}, the results for both modified PC and EMS-PC are presented.
The figures look very similar to Figure ~\ref{fig:struct_avg}, and look strongly symmetric.
In terms of causality, this means that many latent sources are still present and causal relations between the measured brain regions themselves are not clearly visible.

\begin{figure}[h!]
  \centering
  \includegraphics{images/PDAG_avg_mod}
  \caption{The PDAG found with modified PC, in which a high value denotes a high certainty of an arrow. The avarage of all six subjects is taken.}
  \label{fig:pdag_avg_mod}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{images/PDAG_avg_expl}
  \caption{The PDAG found with EMS-PC, in which a high value denotes a high certainty of an arrow. The avarage of all six subjects is taken.}
  \label{fig:pdag_avg_ems}
\end{figure}

Another representation of this data has been used to emphasize the assymetric parts of Figures ~\ref{fig:pdag_avg_mod} and ~\ref{fig:pdag_avg_ems}.
To achieve this, the absolute difference between these matrices and their transpose has been calculated, lifting out those points for which there is an arrow in one direction, but not in the opposite direction.
These results are presented in Figures ~\ref{fig:pdag_avg_antisymmetric_mod} and ~\ref{fig:pdag_avg_antisymmetric_ems}.

\begin{figure}[h!]
  \centering
  \includegraphics{images/PDAG_avg_antisymmetric_mod}
  \caption{The assymetric parts of the PDAG found with modified PC, in which a high value denotes a high certainty of an one-directional arrow. The avarage of all six subjects is taken.}
  \label{fig:pdag_avg_antisymmetric_mod}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics{images/PDAG_avg_antisymmetric_expl}
  \caption{The assymetric parts of the PDAG found with EMS-PC, in which a high value denotes a high certainty of an one-directional arrow. The avarage of all six subjects is taken.}
  \label{fig:pdag_avg_antisymmetric_ems}
\end{figure}

% iets over:
% hoogte van avg_antisymmetric, hoger bij EMS dus meer richting, maar bij beiden niet zo consistent over proefpersonen
% verschillen tussen beide hemispheres: niet zo logisch lijkt me, hemisphers zouden intern ongeveer hetzelfde moeten functioneren
% veel hoge waarden in 'blokjes' rond diagonaal, interessante dingen in de hemispheres?
% per proefpersoon maxima van ~0.5, maar een veel lager avarage over proefpersonen dus weinig consistentie
